Playing Games: A Case Study in Active Learning Applied to Game Theory

### 2015-06-29 paper/paper.tex
Including the wordlike package
Journal requires submission via word.
1st draft of section on Axelrod activity now written.
Various pieces missing.

\usepackage{fullpage} % Package to use full page
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{subcaption}

@@

\begin{figure}
    \begin{center}
        \begin{tikzpicture}
            % First level of play
            \node (0) at (0,0) {$0$};
            \node (100) at (10,0) {$100$};
            \draw [thick] (0) -- (100);
            \node [below] at (6.7,0) {$67$};

@@

\begin{figure}[htpb]
    \begin{subfigure}{.6\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/histogram_of_guesses.pdf}
        \caption{Frequency of guesses depending on the round of play.}
        \label{fig:histogram_of_guess}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/jointplot_of_guesses.pdf}
        \caption{Linear relationship between guesses of each round of play.}
        \label{fig:jointplot_of_guess}
    \end{subfigure}
    \caption{Results from all data collected.}
    \label{fig:all_results}
\end{figure}

@@

\begin{figure}[htpb]
    \begin{subfigure}{.6\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/histogram_of_guesses_removing_increasing_guesses.pdf}
        \caption{Frequency of guesses depending on the round of play.}
        \label{fig:histogram_of_guess}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/jointplot_of_guesses_removing_increasing_guesses.pdf}
        \caption{Linear relationship between guesses of each round of play.}
        \label{fig:jointplot_of_guess}
    \end{subfigure}
    \caption{Results from data when removing increasing guesses.}
    \label{fig:results_with_decreasing_guess}
\end{figure}

@@

This activity is used to introduce students to the concepts of repeated games

\begin{itemize}
    \item The theory
    \item Tournaments:
        \begin{itemize}
            \item Basic type.
            \item Infinitely repeated game.
            \item Markov games.
        \end{itemize}
\end{itemize}

>>>>

\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{subcaption}

\usepackage{wordlike}  % Make this look like word made it

@@

\begin{figure}[!hbtp]
    \begin{center}
        \begin{tikzpicture}
            % First level of play
            \node (0) at (0,0) {$0$};
            \node (100) at (10,0) {$100$};
            \draw [thick] (0) -- (100);
            \node [below] at (6.7,0) {$67$};

@@

\begin{figure}[!hbtp]
    \begin{subfigure}{.6\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/histogram_of_guesses.pdf}
        \caption{Frequency of guesses depending on the round of play.}
        \label{fig:histogram_of_guess}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/jointplot_of_guesses.pdf}
        \caption{Linear relationship between guesses of each round of play.}
        \label{fig:jointplot_of_guess}
    \end{subfigure}
    \caption{Results from all data collected.}
    \label{fig:all_results}
\end{figure}

@@

\begin{figure}[!hbtp]
    \begin{subfigure}{.6\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/histogram_of_guesses_removing_increasing_guesses.pdf}
        \caption{Frequency of guesses depending on the round of play.}
        \label{fig:histogram_of_guess}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/jointplot_of_guesses_removing_increasing_guesses.pdf}
        \caption{Linear relationship between guesses of each round of play.}
        \label{fig:jointplot_of_guess}
    \end{subfigure}
    \caption{Results from data when removing increasing guesses.}
    \label{fig:results_with_decreasing_guess}
\end{figure}

@@

This activity is used to introduce students to the concepts of repeated games
\cite{Maschler2013}. The mathematical details can be omitted from the initial
description of the activity to the participants but for completeness it they are
included here:

A repeated game is played over discrete time periods. Each time period is
indexed by \(0<t\leq T\) where \(T\) is the total number of periods.  In each
period \(N\) players play a static game referred to as the \textbf{stage game}
independently and simultaneously selecting actions.  Players make decisions in
full knowledge of the \textbf{history} of the game played so far (ie the actions
chosen by each player in each previous time period).  The payoff is defined
as the sum of the utilities in each stage game for every time period.

One of the most renowned repeated games is referred to as \textbf{Axelrod's
tournament} which is what is recreated in this activity \cite{Axelrod1980a,
Axelrod1980b}.

Initially a description of the prisoner's dilemma \cite{Maschler2013} is given.
The prisoner's dilemma is a simple two player game that is often used to
introduce the very basic notions of game theory. It is described by the
following two matrices:

\[
    A =
    \begin{pmatrix}
        3&5\\
        0&1
    \end{pmatrix}
    \qquad
    B =
    \begin{pmatrix}
        3&0\\
        5&1
    \end{pmatrix}
\]

The row player has utility given by \(A\) and the column player has utilities
given by \(B\).
The strategies available to each player are to cooperate: \(C\) or to defect:
\(D\). Playing \(C\) corresponds to players choosing their first row/column and
\(D\), the second row/column.

Thus if both players cooperate they both receive a utility of 3, if one player
defects, the defector gets a utility of 5 and the cooperator a utility of 0.
Finally if both players defect they receive a utility of 1. As players (in this
framework) aim to maximise their score, the Nash equilibrium for this game is
for both players to defect.

After describing this activity and in particular explaining the simple
mathematical idea of \textbf{dominated strategy} (which is what is used in the
activity of Section \ref{sec:best_responses}) participants are made aware of the
concept of Nash equilibrium. This in turn can lead to a brief description of the
tragic yet brilliant life of John Nash.

At this point the activity is described:

\begin{enumerate}
    \item All participants will form four groups/teams;
    \item Teams will \textit{duel} each other in repetitions of 5 to 8 rounds
        (depending on available time).
    \item All teams will play in a round robin tournament with cumulative scores
        being recorded.
    \item The victorious team will be the team with the highest total score.
\end{enumerate}

The tournament is run with all participants present (even those not watching).
All participants are invited to stand and confer in their teams. The importance
of standing (as a physical activity) is noted in \cite{Donnelly2011} (whilst
that reference is mainly concerned with the impact of activity on physical well
being it also describes advantages in terms of concentration).  Before every
round of every duel opposing teams are encouraged to discuss strategies, after
which they do not face each other and following a prompt hold up a card
indicating either \(C\) or \(D\).  Duels are recorded on the white board in a
table similar to the one shown in Table \ref{tab:duel}, in which two strategies
constantly cooperate (thus obtaining a utility of 3 in each round). Table
\ref{tab:duel_1} shows an example where a strategy that is alternating plays
against a strategy that always defects.

\begin{table}[!htbp]
    \caption{Playing Tit For Tat against Cooperator}
    \centering
    \begin{tabular}{ccccc}
        \toprule
        3&6&9&12&15\\
        \midrule
        3&6&9&12&15\\
        \bottomrule
    \end{tabular}
    \label{tab:duel}
\end{table}

\begin{table}[!htbp]
    \caption{Playing Alternator against Defector}
    \centering
    \begin{tabular}{ccccc}
        \toprule
        5&6&11&12&17\\
        \midrule
        0&1&1&2&2\\
        \bottomrule
    \end{tabular}
    \label{tab:duel_1}
\end{table}

The names of the strategies shown in Tables \ref{tab:duel} and \ref{tab:duel_1}
are strategies that were used in the original tournaments run by Axelrod
\cite{Axelrod1980a, Axelrod1980b}. The interesting fact of repeated games and
one that becomes apparent to participants through the activity is that whilst
repeating the stage Nash equilibrium (always defect) is indeed a Nash
equilibrium for the repeated games, this equilibrium is not unique as
reputation now has a part to play.

Note that if students do not realise this it is important to remind them that
the goal is not to win each duel but to obtain a high score. Often during the
tournament one team will (during the pre round discussion) exclaim:

\begin{quote}
   ``We will cooperate until you defect, at which point we will defect
throughout''
\end{quote}

Without realising it the participants have described a well known strategy
(\textbf{Grudger}) which takes in to account the entire history of play.
Figure \ref{fig:} shows a photo of a final board for a particular implementation
of this activity.

This activity can be complemented with a demonstration of software that allows
for the rapid simulation of Axelrod's tournament \cite{Knight2015}. Figure
\ref{fig:evolutionary_axelrod} shows the performance of the strategies when put
in an evolutionary context.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=.6\textwidth]{static/basic_strategies-reproduce.pdf}
    \caption{Repeated games in an evolutionary context}
    \label{fig:evolutionary_axelrod}
\end{figure}

Weakness about all participants observing...

As well as being enjoyable this activity is a good intro to...

Furthermore can be modified to study ...
